\documentclass{beamer}
\mode<presentation>{\usetheme{Dresden}}

\usepackage{fancyvrb}

\title{High-Level Program Execution\\ for Multi-Agent Teams}
\author{Ryan Kelly}

\begin{document}


\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

\section{Motivation}

\begin{frame}
\frametitle{Multi-Agent Teams}
Two categories of multi-agent system:
\begin{itemize}
  \item Open multi-agent systems
  \item Multi-agent teams
\end{itemize}
\ \\
\ \\
Teams may be conceptualised as a single agent, with distributed sensing,
reasoning and acting capabilities.\\
\ \\
\pause
What ideas can we borrow from single-agent programming?
\end{frame}

\begin{frame}
\frametitle{High-Level Program Execution}
A practical alternative to planning for single-agent systems. Specify
a program made up of:
\begin{itemize}
  \item Actions from the world
  \item Connected by high-level programming constructs
  \item Including nondeterminism where reasoning is required
\end{itemize}
\ \\
\ \\
Why?
\begin{itemize}
  \item Controlled Nondeterminism
  \item Clear, simple programs
  \item Symbolic reasoning about the world
  \item Can prove program properties (safety/liveness/etc...)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Motivating Example: The Cooking Agents}
Several robotic chefs inhabit a kitchen, along with various ingredients,
appliances and utensils.  They must cooperate to produce a meal consisting
of several dishes.\\
\ \\
Challenges:
\begin{itemize}
  \item Concurrent execution of tasks
  \item Conflict over shared resources
  \item Coordination of shared actions
\end{itemize}

Assumptions:
\begin{itemize}
  \item Deterministic, fully observable world
  \item Complete and reliable communication
\end{itemize}
\end{frame}


\section{Golog Review}

\begin{frame}
\frametitle{The Situation Calculus}

A first-order logic formalism for reasoning about dynamic worlds:

\begin{itemize}
\pause
\item Actions: $wait$, $pickup(obj)$, $goto(x,y)$
\pause
\item Situations: $S_{0}$, $do(a_{1},S_{0})$, $do(a_{2},do(a_{1},S_{0}))$
\pause
\item Fluents: $holding(obj,s)$, $(x,y)=curPosition(s)$
\pause
\item Precondition Axioms: $Poss(pickup(obj),s) \leftrightarrow holding(obj,s)$
\pause
\item Successor State Axioms:\[
\begin{array}{cc}
holding(obj,do(a,s))\iff & a=pickup(obj)\,\vee\\
 & \left[holding(obj,s)\wedge a\neq drop(obj)\right]\end{array}\]
\end{itemize}
\pause
Successor state axioms can provide an elegant solution to the infamous
frame problem.

\end{frame}

\begin{frame}
\frametitle{Golog}
Introduces programming to the S.C. by means of "complex actions":
\begin{itemize}
  \pause
  \item $\delta_1;\delta_2$: Perform actions in sequence
  \pause
  \item $\phi?$: Assert that a condition holds
  \pause
  \item $\delta_1|\delta_2$: Choose between actions to perform
  \pause
  \item $(\pi x)\delta(x)$: Choose suitable bindings for variables
  \pause
  \item $\delta^*$: Perform an action zero or more times
  \pause
\end{itemize}
Composing actions by means of these operators allows complex programs
to be defined.
\end{frame}

\begin{frame}
\frametitle{Example: A Salad in Golog}
Consider a Golog program for making a simple salad:\\
\ \\
\begin{columns}
  \begin{column}{0.5\textwidth}
\[
\begin{array}{c}
\mathbf{proc}\ ChopTypeInto(type,dest)\\
\pi(obj)[IsType(obj,type)?\ ;\\
chop(obj)\ ;\ place\_in(obj,dest)]\\
\mathbf{end}\end{array}\]
  \end{column}
  \begin{column}{0.5\textwidth}
\[
\begin{array}{c}
\mathbf{proc}\ MakeSalad(dest)\\
ChopTypeInto(Lettuce,dest)\ ;\\
ChopTypeInto(Carrot,dest)\ ;\\
ChopTypeInto(Tomato,dest)\ ;\\
mix(dest,1)\\
\mathbf{end}\end{array}\]
  \end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Executing a Golog Program}
Intuitively, an agent must find a \emph{legal execution} of its nondeterministic
program - some set of choices for the nondeterministic components so that all
actions performed are possible, and all test conditions are satisfied.\\
\ \\
Formally, the agent must find a situation $s$ such that:\[
\mathbf{D} \models \exists s . Do(\delta,S_0,s)\]

Here, $\mathbf{D}$ is the Situation Calculus theory of action, and $Do$
encodes the semantics of the Golog operators.  Remember, a situation is a sequence of actions - in this case, the actions which must be carried out to execute the program.
\end{frame}

\begin{frame}
\frametitle{ConGolog: Concurrency and Interrupts}
"Concurrent Golog" was designed to support the concurrent execution of
several threads of control.  It also allows threads to be spawned in
response to conditions in the environment: \\
\ \\
\begin{itemize}
  \pause
  \item $\delta_1\ ||\ \delta_2$: Execute two programs concurrently
  \pause
  \item $\delta_1\ \gg\ \delta_2$: Execute two programs concurrently, giving one priority over the other
  \pause
  \item $<\phi \rightarrow \delta>$: Execute $\delta$ when $\phi$ becomes true
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{ConGolog}
Semantics are defined in terms of "single steps" of computation.
$Trans(\delta,s,\delta',s')$ indicates that program $\delta$ may execute
a single step, taking the world from situation $s$ to $s'$, and leaving the
program $\delta'$ still to be executed.\\
\ \\
\pause
Concurrency is achieved by interleaving steps from both programs:
\[
\begin{array}{cc}
Trans(\delta_1\ ||\ \delta_2,s,\delta',s') \equiv & \exists \gamma . \delta'=(\gamma\ ||\ \delta_2)\wedge Trans(\delta_1,s,\gamma,s')\\
& \vee\ \exists \gamma . \delta'=(\delta_1\ ||\ \gamma)\wedge Trans(\delta_2,s,\gamma,s')
\end{array}\]
\end{frame}

\begin{frame}
\frametitle{A Salad in ConGolog}
When making a salad, the order of ingredients doesnt matter.  This can
be expressed using concurrency:\[
\begin{array}{c}
\mathbf{proc}\ MakeSalad(dest)\\
\left[\ ChopTypeInto(Lettuce,dest)\ ||\right.\\
ChopTypeInto(Carrot,dest)\ ||\\
\left.ChopTypeInto(Tomato,dest)\ \right]\ ;\\
mix(dest,1)\\
\mathbf{end}\end{array}\]
\end{frame}


\begin{frame}
\frametitle{IndiGolog: Online Execution and Sensing}
In order to be sure an execution is legal, Golog must plan all the way
to the end.  "Offline Execution" - for highly nondeterministic programs
this may be intractable.\\
\ \\
Furthermore, the agent might not have all the information it needs when
execution begins - may require \emph{sensing}.

\begin{itemize}
  \pause
  \item Execute online, making nondet choices arbitrarily
  \pause
  \item Perform sensing actions when extra information is required
  \pause
  \item Allow planning over individual parts of the program: $\Sigma(\delta)$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A Salad in IndiGolog}
Regardles of what ends up in the bowl, it can always be mixed.  Only need
to plan how to add the ingredients:\[
\begin{array}{c}
\mathbf{proc}\ MakeSalad(dest)\\
\Sigma\left[\ ChopTypeInto(Lettuce,dest)\ ||\right.\\
ChopTypeInto(Carrot,dest)\ ||\\
\left.ChopTypeInto(Tomato,dest)\ \right]\ ;\\
mix(dest,1)\\
\mathbf{end}\end{array}\]
\end{frame}


\section{Golog for MA Teams?}

\begin{frame}
\frametitle{Golog for MA Teams?}
Golog has proven very powerful in single-agent domains.  But, there are
some challenges when moving to a multi-agent setting:
\begin{itemize}
  \pause
  \item Several actions can be performed at the same instant
  \pause
  \item Need to coordinate concurrent actions
  \pause
  \item Need to predict actions of others, and the environment
  \pause
  \item Agents may have differing knowledge/belief about the world
\end{itemize}
\ \\
\ \\
\pause
Some of these have been individually addressed in the Situation Calculus.
We integrate them into a Golog suitable for multi-agent teams: "MIndiGolog".

\end{frame}

\section{Concurrency}

\begin{frame}
\frametitle{True vs Interleaved Concurrency}
ConGolog provides \emph{interleaved concurrency} allowing multiple
high-level programs to be executed together.\\
\ \\
With mulitple agents, the world exhibits \emph{true concurrency}. Several
actions can occur at the same instant.\\
\ \\
\pause
Need to:
\begin{itemize}
  \item allow true concurrency to be represented
  \item modify the $||$ operator to take advantages of true concurrency
  \item ensure that this modification is well-behaved
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Concurrent Situation Calculus (Reiter, 1996)}
Have the first argument of an action name the agent performing it:\[
pickup(Thomas,Lettuce1),\ mix(Richard,Bowl1,1)\]
\pause
Replace action terms with sets of actions, all of which are performed
at the same time:\[
do(\{pickup(Thomas,Lettuce1),pickup(Richard,Tomato1)\},S_0)\]
\pause
Precondition interaction:\[
Poss(\{pickup(Richard,Obj),pickup(Harriet,Obj)\},s)\ ?\]
\pause
Basic solution:\[
Poss(C,s) \leftrightarrow \forall a \in C.Poss(a,s)\ \wedge\ \neg Conflicts(C,s)\]
\end{frame}

\begin{frame}
\frametitle{True + Interleaved Concurrency}
When executing two programs in parallel, it may be possible for the team to
execute a step from each simultaneously:\[
\begin{array}{cc}
Trans(\delta_1\ ||\ \delta_2,s,\delta',s') \equiv & \exists \gamma . \delta'=(\gamma\ ||\ \delta_2)\wedge Trans(\delta_1,s,\gamma,s')\\
& \\
& \vee\ \exists \gamma . \delta'=(\delta_1\ ||\ \gamma)\wedge Trans(\delta_2,s,\gamma,s')\\
& \\
& \vee\ \exists c_1,c_2,\gamma_1,\gamma_2.\ Trans(\delta_1,s,\gamma_1,do(c_1,s)\\
& \wedge\ Trans(\delta_2,s,\gamma_2,do(c_2,s))\\
& \wedge\ \delta'=(\gamma_1||\gamma_2) \wedge s'=do(c_1 \cup c_2,s)\\
\end{array}\]
But, this simple encoding is not well-behaved (later...)
\end{frame}

\section{Time}

\begin{frame}
\frametitle{Time}
Which is easier for a team to manage?
\begin{itemize}
  \pause
  \item Perform a list of actions concurrently
  \pause
  \item Perform a list of actions concurrently at time $t$
\end{itemize}
\ \\
\ \\
\pause
Assuming a good wall-clock is available, explicit time makes coordination easier.\\
\ \\
It is also generally a nice feature to have when representing dynamic worlds.
Example: bake a cake for \emph{10 minutes}
\end{frame}

\begin{frame}
\frametitle{The Temporal Situation Calculus}
Adding time to the S.C. has a long history (Pinto 1994, Reiter 1996).\\
\ \\
We opt for a simple solution: attach an explicit time to each situation,
indicating when it was brought about:\[
\begin{array}{c}
do(a_1,t_1,S_0),\ do(a_2,t_2,do(a_1,t_1,S_0))\\
Poss(a,t,s) \leftrightarrow\ \dots\\
start(do(a,t,s)) = t
\end{array}\]
\pause
Temporal Golog finds situations with \emph{constraints} on the
time at which actions are performed:\[
\begin{array}{c}
do(a_2,t_2,do(a_1,t_1,S_0)\\
t_1>0,\ t_2>t_1+5
\end{array}\]
\end{frame}

\begin{frame}
\frametitle{Continuous Actions}
Some actions have a finite duration: $mix(Thomas,Bowl1,5)$
Others are instantaneous: $pickup(Robert,Lettuce1)$.
So this concurrent action is problematic:\[
\{mix(Thomas,Bowl1,5),pickup(Robert,Lettuce1)\}\]
Standard solution: decompose continuous actions into instantaneous $begin$ 
and $end$ actions:\[
begin\_mix(Thomas,Bowl1,5) \dots end\_mix(Thomas,Bowl1,5)\]
But $end\_mix()$ is not a standard action.  It cannot be performed at
any time.  Rather, it \emph{must} occur whenever it is possible.
\end{frame}

\section{Natural Actions}

\begin{frame}
\frametitle{Natural Actions (Reiter, 1996)}
\begin{itemize}
  \item Natural Action:  an action that must occur whenever it is possible,
        unless some earlier action prevents it from occuring.
  \item $Poss(na,t,s)$ predicts the times at which a natural action will occur.
  \item Least Natural Time Point: the earliest time at which a natural
        action will occur for a given situation:\[
\begin{array}{c}
lntp(s,t) \leftrightarrow \exists a [natural(a) \wedge Poss(a,t,s)] \wedge\\
\forall a,t_a [natural(a) \wedge Poss(a,t_a,s) \rightarrow t \leq t_a]
\end{array}\]
  \item Legal Situation: one in which any natural actions that could have occured, did occur:\[
\begin{array}{c}
legal(do(c,t,s)) \leftrightarrow legal(s) \wedge\\
Poss(c,t,s) \wedge start(s) \leq t \wedge\\
\forall a,t_a [natural(a) \wedge Poss(a,t_a,s) \rightarrow (a \in c \vee t < t_a)]
\end{array}\]
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our Semantics}
Most ConGolog/IndiGolog operators trivially modified for the extended S.C.\\
\ \\
\pause
$Trans(c,s,\delta',s')$ modified to ensure only legal situations are produced
(too complicated for this slide)
\pause
\[\begin{array}{cc}
Trans(\delta_1\ ||\ \delta_2,s,\delta',s') & \equiv \exists \gamma . \delta'=(\gamma\ ||\ \delta_2)\wedge Trans(\delta_1,s,\gamma,s')\\
& \vee\ \exists \gamma . \delta'=(\delta_1\ ||\ \gamma)\wedge Trans(\delta_2,s,\gamma,s')\\
& \vee\ \exists c_1,c_2,\gamma_1,\gamma_2,t.\ Trans(\delta_1,s,\gamma_1,do(c_1,t,s)\\
& \wedge\ Trans(\delta_2,s,\gamma_2,do(c_2,t,s)) \wedge \delta'=(\gamma_1||\gamma_2)\\
& \wedge \forall a\left[a \in c_1 \wedge a \in c_2 \rightarrow natural(a) \right]\\
& \wedge\ Poss(c_1 \cup c_2,t,s) \wedge s'=do(c_1 \cup c_2,t,s)\\
\end{array}\]
\end{frame}

\begin{frame}
\frametitle{Making Salad, again}
So, how does a team cooperate to make a salad?
\[
\begin{array}{c}
\mathbf{proc}\, MakeSalad(dest)\\
\left[\pi(agt)(ChopTypeInto(agt,Lettuce,dest))\,||\right.\\
\pi(agt)(ChopTypeInto(agt,Carrot,dest))\,||\\
\left.\pi(agt)(ChopTypeInto(agt,Tomato,dest))\right]\,;\\
\pi(agt)\left[acquire(agt,dest)\,;\right.\\
\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, begin\_ task(agt,mix(dest,1))\,;\\
\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, end\_ task(agt,mix(dest,1))\,;\\
\left.\,\,\,\, release(agt,dest)\right]\,\,\mathbf{end}\end{array}\]
\end{frame}

\section{Implementation}

\begin{frame}
\frametitle{Implementation 1}
Our first implementation was in Prolog:
\begin{itemize}
  \item $Trans$ rules from the semantics of ConGolog/IndiGolog
  \pause
  \item Modified for lists of concurrent actions, explicit time
  \pause
  \item Including our new rules for $||$ and primitive actions
  \pause
  \item And an axiomatisation of the "cooking agents" domain
\end{itemize}
\ \\
\ \\
\pause
It functions as a \emph{centralised} multi-agent planner.
\end{frame}

\begin{frame}
\frametitle{Actually Making Salad}
\begin{columns}
  \begin{column}{0.5\textwidth}
    \VerbatimInput[fontsize=\tiny]{output_makeSalad_1.txt}
  \end{column}
  \begin{column}{0.5\textwidth}
    \VerbatimInput[fontsize=\tiny]{output_makeSalad_2.txt}
  \end{column}
\end{columns}
\ \\
\ \\
Can give idle agents other work using:\[
MakeSalad(Bowl1)\ ||\ MakeCake(Bowl2)\ ||\ \dots\]
\end{frame}

\begin{frame}
\frametitle{Implementation 2}
But, we want to take advantage of the distributed reasoning capabilities of
the team.  Since we have a logic-based language, we need
\emph{Distributed Logic Programming}.\\
\ \\
Oz is a multi-paradigm language with strong support for distributed execution.
In particular, it can do parallel search over logic programs:\\
\ \\
\VerbatimInput[fontsize=\scriptsize]{oz_code.txt}
\ \\
This version can function as a \emph{distributed} multi-agent planner
\end{frame}

\section{Related/Future Work}

\begin{frame}
\frametitle{Related Work}
\emph{Many} extensions to the situation calculus and Golog.  We were able to
pick and choose the ones we needed.\\
\ \\
Multi-agent Golog:
\begin{itemize}
  \item CASL: Modelling and Simulation of Multi-Agent Systems (Levesque, 2002-)
  \item ReadyLog: Golog for Team Coordination and Control (Lakemeyer, 2002-)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Future Work}
So far, agents are assumed to be fully informed, and the world
fully obersvable.  This is not realistic in many situations. 
Investigating:
\begin{itemize}
  \item Knowledge and Belief in the S.C.
  \item Sensing from IndiGolog
  \item Modal logics of Knowledge and Belief
\end{itemize}
\ \\
\ \\
While planning uses the teams' distributed resources, control of the
process is still centralised.  How can the agents plan together without
centralised control?
\end{frame}

\end{document}
