\documentclass{beamer}

\newcommand{\isdef}{\hbox{$\stackrel{\mbox{\tiny def}}{=}$}}
\newcommand{\Dt}{\mathcal{D}}
\newcommand{\Reg}{\mathcal{R}}
\newcommand{\Pst}{\mathcal{P}}
\newcommand{\Trn}{\mathcal{T}}
\newcommand{\Kln}{\mathcal{K}}
\newcommand{\TrnA}{\Trn_{a}}
\newcommand{\EKnows}{\mathbf{EKnows}}
\newcommand{\Knows}{\mathbf{Knows}}
\newcommand{\CKnows}{\mathbf{CKnows}}
\newcommand{\KnowsZ}{\mathbf{Knows_{0}}}
\newcommand{\PKnowsZ}{\mathbf{PKnows_{0}}}
\newcommand{\PKnows}{\mathbf{PKnows}}
\newcommand{\KTrans}{\mathbf{KDo}}
\newcommand{\KDo}{\mathbf{KDo}}
\newcommand{\EDo}{\mathbf{EDo}}
\newcommand{\vars}[1]{\bar{#1}}
\newcommand{\PbU}{PbU}

\mode<presentation>{\usetheme{Dresden}}

\usepackage{fancyvrb}

\title{Asynchronous Multi-Agent Domains\\ in the Situation Calculus}
\author{Ryan Kelly\\ (Supervisor: Adrian Pearce)}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \begin{itemize}
  \item Situation Calculus
  \item MIndiGolog
  \item Observations
  \item Knowledge
  \item Common Knowledge
  \item Joint Execution
  \item Conclusions
  \end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Situation Calculus}
\emph{Actions} are instantaneous events causing the world to change
\begin{itemize}
  \item $pickup(Thomas,Bowl)$, $beginTask(Richard,mix(Bowl,1))$
\end{itemize}
\emph{Situations} are histories of actions that have been performed
\begin{itemize}
  \item $S_0$, $do(pickup(Harriet,Knife),S_0)$
\end{itemize}
\emph{Fluents} are situation-dependent properties of the world
\begin{itemize}
  \item $Poss(a,s)$, $Holding(Harriet,Knife,s)$
\end{itemize}
\emph{Successor State Axioms} define what holds after an action in terms of what held before the action
\begin{itemize}
  \item $Holding(agt,obj,do(a,s)) \equiv a = pickup(agt,obj)$ \\
        $\,\,\,\,\,\,\,\,\,\,\,\,\,\vee Holding(agt,obj,s) \wedge \neg\left(a = drop(agt,obj)\right)$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Situation Calculus}
From these components we can built a first-order-logic \emph{Theory of Action}
$\mathbf{D}$, with which agents can:
\begin{itemize}
  \item Reason about when actions are possible
  \item Reason about the effects of actions
  \item Represent sequences of actions as situations:\[
\left[a_1, a_2, a_3, a_4\right] \rightarrow do(a_4,do(a_3,do(a_2,do(a_1,S_0))))
\]
\end{itemize}
\pause
But, what about more complex actions and procedures?
\begin{itemize}
  \item $\mathbf{if}\ available(Item)\ \mathbf{then}\ pickup(Item)\ \mathbf{else}\ beSad$
  \item $\mathbf{while}\ (\exists block)\ ontable(block)\ \mathbf{do}\ remove(block)\ \mathbf{endWhile}$
  \item $\mathbf{proc}\ remove(block)\ [pickup(block);putaway(block)]\ \mathbf{endProc}$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{GOLOG: "alGOL in LOGic"}
GOLOG introduces programs to the SitCalc via "complex actions".
Primitive actions are composed using operators such as:

\begin{itemize}
  \pause
  \item $\delta_1;\delta_2$ - Perform two programs in sequence
  \pause
  \item $\phi?$ - Assert that a condition holds
  \pause
  \item $\delta_1|\delta_2$ - Choose between programs to execute
  \pause
  \item $(\pi x)\delta(x)$ - Choose suitable bindings for variables
  \pause
  \item $\delta^*$ - Execute a program zero or more times
\end{itemize}
\ \\
\ \\
\pause
Key Point:  programs can include \alert{nondeterminism}
\end{frame}

\begin{frame}
\frametitle{A Simple Example}
Consider a Golog program for getting to uni of a morning:\[
\begin{array}{c}
ringAlarm;(hitSnooze; ringAlarm)^*;turnOffAlarm;\\
(\pi\ food)(edible(food)?;eat(food)); haveShower; brushTeeth\\
(driveToUni\ |\ trainToUni); (time<9:00)?
\end{array}\]

There are potentially many legal executions, depending on which actions
are possible in the world.

Use the theory of action to inform nondeterministic choices.
\end{frame}

\begin{frame}
\frametitle{Why the Situation Calculus?}

\end{frame}

\section{MIndiGolog}

\begin{frame}
\frametitle{Golog for Multi-Agent Teams?}
Is Golog a useful formalism for multi-agent teams?
Consider programming our cooking agents to make a simple salad:
\ \\
\begin{columns}
  \begin{column}{0.5\textwidth}
\[
\begin{array}{c}
\mathbf{proc}\ ChopTypeInto(type,dest)\\
\pi(obj)[IsType(obj,type)?\ ;\\
chop(obj)\ ;\ place\_in(obj,dest)]\\
\mathbf{end}\end{array}\]
  \end{column}
  \begin{column}{0.5\textwidth}
\[
\begin{array}{c}
\mathbf{proc}\ MakeSalad(dest)\\
ChopTypeInto(Lettuce,dest)\ ;\\
ChopTypeInto(Carrot,dest)\ ;\\
ChopTypeInto(Tomato,dest)\ ;\\
mix(dest,1)\\
\mathbf{end}\end{array}\]
  \end{column}
\end{columns}
\ \\
\ \\
\pause
\alert{Idea}:  have the team cooperate to execute a shared Golog program
\end{frame}

\begin{frame}
\frametitle{Multi-Agent Situation Calculus}
What do we need for a multi-agent situation calculus?
\begin{itemize}
  \item Identify which agent performs which action
  \item True concurrency of actions
  \item Explicit account of time, for coordination
  \item Natural actions, the predict the behavior of the world
  \item Knowledge about state of the world
  \item Knowledge about what actions have been performed
\end{itemize}
\ \\
\ \\
\pause
Many of these have been addressed independently by other authors.

We extend and combine them to form a compelling formalism for multi-agent
worlds.
\end{frame}

\begin{frame}
\frametitle{Multiple Agents, Explicit Time}
Multiple agents can be handled by having the first argument of each action
identify the agent performing it.
\begin{itemize}
  \item $pickup(Richard,Block1)$
  \item $goto(Thomas,x,y)$
  \item $chop(Harriet,Tomato1)$
\end{itemize}
\pause
Adding time to the SitCalc has a long history \cite{reiter01kia}.
We opt for a simple solution: attach an explicit time to each situation,
indicating when it was brought about:\[
\begin{array}{c}
do(a_1,t_1,S_0),\ do(a_2,t_2,do(a_1,t_1,S_0))\\
Poss(a,t,s) \equiv\ \dots\\
start(do(a,t,s)) = t
\end{array}\]
\end{frame}

\begin{frame}
\frametitle{True vs Interleaved Concurrency}
ConGolog provides \emph{interleaved concurrency} allowing multiple
high-level programs to be executed together.\\
\ \\
With multiple agents, the world exhibits \emph{true concurrency}. Several
actions can occur at the same instant.\\
\ \\
\pause
Need to:
\begin{itemize}
  \item allow true concurrency to be represented
  \item modify the $||$ operator to take advantage of true concurrency
  \item ensure that this modification is well-behaved
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Concurrent Situation Calculus}
Replace action terms with sets of actions, all of which are performed
at the same time \cite{reiter01kia}:\[
do(\{pickup(Thomas,Lettuce1),pickup(Richard,Tomato1)\},t_1,S_0)\]
\pause
\ \\
\ \\
Precondition interaction:\[
Poss(\{pickup(Richard,Obj),pickup(Harriet,Obj)\},t,s)\ ?\]
\pause
Basic solution:\[
Poss(c,s) \equiv \forall a \in c.Poss(a,s)\ \wedge\ \neg Conflicts(c,s)\]
\end{frame}

\begin{frame}
\frametitle{Continuous Actions}
Some actions have a finite duration: $mix(Thomas,Bowl1,5)$
Others are instantaneous: $pickup(Robert,Lettuce1)$.
So this concurrent action is problematic:\[
\{mix(Thomas,Bowl1,5),pickup(Robert,Lettuce1)\}\]
Standard solution: decompose continuous actions into instantaneous $begin$
and $end$ actions:\[
begin\_mix(Thomas,Bowl1,5) \dots end\_mix(Thomas,Bowl1,5)\]
But $end\_mix()$ is not a standard action.  It cannot be performed at
any time.  Rather, it \emph{must} occur whenever it is possible.
\end{frame}

\begin{frame}
\frametitle{Natural Actions}
Natural Action:  an action that must occur whenever it is possible,
unless some earlier action prevents it from occurring.
\begin{itemize}
  \item $Poss(na,t,s)$ predicts the times at which a natural action will occur.
  \item Legal Situation: one in which any natural actions that could have occurred, did occur:\[
\begin{array}{c}
legal(do(c,t,s)) \equiv legal(s) \wedge\\
Poss(c,t,s) \wedge start(s) \leq t \wedge\\
\forall a,t_a [natural(a) \wedge Poss(a,t_a,s) \rightarrow (a \in c \vee t < t_a)]
\end{array}\]
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{MIndiGolog}
We have created a new Golog variant that:
\begin{itemize}
  \item Integrates both true and interleaved concurrency
  \item Has an explicit notion of time
  \item Automatically handles natural actions
\end{itemize}
\ \\
\ \\
Implemented in Prolog, it functions as a \emph{centralised} multi-agent planner.
\ \\
\ \\
Using the distributed logic programming capabilities of Oz, the planning
workload can be \emph{distributed} amongst the agents.
\end{frame}

\begin{frame}
\frametitle{Making a Salad}
When making a salad, the order of ingredients doesn't matter.  This can
be expressed using concurrency:\[
\begin{array}{c}
\mathbf{proc}\ MakeSalad(dest)\\
\left[\ ChopTypeInto(Lettuce,dest)\ ||\right.\\
ChopTypeInto(Carrot,dest)\ ||\\
\left.ChopTypeInto(Tomato,dest)\ \right]\ ;\\
mix(dest,1)\\
\mathbf{end}\end{array}\]
We must also expand the continuous $mix$ action, and specify the agent
performing each action.
\end{frame}

\begin{frame}
\frametitle{Making a Salad}
But, it doesn't matter who does what: more nondeterminism
\[
\begin{array}{c}
\mathbf{proc}\, MakeSalad(dest)\\
\left[\pi(agt)(ChopTypeInto(agt,Lettuce,dest))\,||\right.\\
\pi(agt)(ChopTypeInto(agt,Carrot,dest))\,||\\
\left.\pi(agt)(ChopTypeInto(agt,Tomato,dest))\right]\,;\\
\pi(agt)\left[begin\_task(agt,mix(dest,1))\,;\right.\\
\left.\,\,\,\, end\_task(agt,mix(dest,1))\right]\,\,\mathbf{end}\end{array}\]
\pause
Of course, we could hide this with syntactic sugar...
\end{frame}

\begin{frame}
\frametitle{Actually Making Salad}
\begin{columns}
  \begin{column}{0.5\textwidth}
    \VerbatimInput[fontsize=\tiny]{output_makeSalad_1.txt}
  \end{column}
  \begin{column}{0.5\textwidth}
    \VerbatimInput[fontsize=\tiny]{output_makeSalad_2.txt}
  \end{column}
\end{columns}
\ \\
\ \\
\pause
Can give idle agents other work using:\[
MakeSalad(Bowl1)\ ||\ MakeCake(Bowl2)\ ||\ \dots\]
\end{frame}

\begin{frame}
\frametitle{Golog for Multi-Agent Teams!}
We get all the benefits of Golog for single-agent systems:
\begin{itemize}
  \item Controlled Nondeterminism
  \item Clear, simple programs
  \item Symbolic reasoning about the world
  \item Can prove safety/liveness/etc properties
\end{itemize}
\ \\
\pause
Plus:
\begin{itemize}
  \item The same program can be given to \emph{any} team of agents
  \item Conflicts and concurrency handled declaratively
  \item Planning workload easily shared between agents
\end{itemize}
\ \\
\pause
Golog has been successfully used for a RoboCup team \cite{Ferrein2005readylog}.
\end{frame}


\section{Ongoing Work}

\begin{frame}
\frametitle{Ongoing Work}

The current implementation assumes the world is fully observable.
\begin{itemize}
  \item What if agents have different ideas about the state of the world?
  \item What is they can't see what the other agents are doing?
\end{itemize}
\ \\
\ \\
Although the workload can be distributed, \emph{control} over planning 
an execution still centralised.

\begin{itemize}
  \item How can the execution be decentralised?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Knowledge}
Knowledge in the S.C. is based on the standard "possible worlds" model:
\[
\begin{array}{c}
\mathbf{Knows}(agt,\phi,s)\ \isdef\ \forall s'\ K(agt,s',s) \rightarrow \phi(s')\\
\\
K(agt,s'',do(a,s)) \equiv \exists s'\ s''=do(a,s') \wedge K(agt,s',s)
\end{array}
\]
\pause
This assumes that all actions are public.  This is unreasonable
in a multi-agent setting, so we explicitly axiomatise action observability:
\[
\begin{array}{c}
Unobs(agt,s'',s)\ \isdef\ \neg\exists s',a\ s<do(a,s')\leq s'' \wedge CanObs(agt,a,s')\\
\\
K(agt,s'',do(a,s)) \equiv \exists s'\ Unobs(agt,do(a,s'),s'') \wedge K(agt,s',s)
\end{array}
\]

\end{frame}

\begin{frame}
\frametitle{Reasoning about Knowledge}
Our account of knowledge requires reasoning about all possible future
situations.  This typically requires second-order reasoning (Uh-Oh!)
\pause
\ \\
\ \\
We have developed an algorithm to calculate the "persistence conditions"
of a formula - the conditions under which a formula is guaranteed to
remain true as long as a certain class of actions do not occur:
\[
\mathcal{P}[\phi,\alpha] \equiv \forall s'' . s \leq s'' \rightarrow \phi(s'') \vee \exists a,s'\ s<do(a,s')\leq s'' \wedge \alpha(a,s)
\]
\pause
We can thus handle multi-agent knowledge using purely first-order reasoning
(Phew!)
\end{frame}

\section{Conclusions}

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
  \item High-Level Program Execution is a promising paradigm for multi-agent teams
\end{itemize}
\pause
Contributions so far:
\begin{itemize}
  \item Expanded Golog semantics to handle true concurrency, natural actions
  \item Developed more general theory of multi-agent knowledge
  \item Identified "tractable" reasoning procedure for knowledge
\end{itemize}
\pause
Ongoing:
\begin{itemize}
  \item Developing knowledge-based distributed coordination strategy
  \item Building a reusable M.A. programming system\\(Prolog? Oz? Mercury?)
\end{itemize}
\end{frame}

\bibliography{/storage/uni/pgrad/library/references}
\bibliographystyle{plain}

\end{document}
