

\chapter{Available Software Implementations}

\label{ch:implementation}

Three major software components have been developed during the course
of this research. They are each made available under the terms of
the GNU General Public License, and are available for download at
the author's website:

\begin{center}
\url{http://www.rfk.id.au/research/thesis/} 
\par\end{center}

Each software package comes with comprehensive instructions on running
the code and reproducing the results found in this thesis. The following
is a brief description of each system.

\begin{description}
\item [{{MIndiGolog~v1:}}] This is the MIndiGolog implementation described
in Chapter \ref{ch:mindigolog}, which performs online execution planning
but is limited to synchronous domains. This software runs on the Mozart
platform version 1.3.2 or later. Its key feature is the use of Mozart's
parallel search functionality to distribute the execution planning
workload. 
\item [{{MIndiGolog~v2:}}] This is the MIndiGolog implementation described
in Chapter \ref{ch:jointexec}, which produces joint executions as
the output of its planning process, but is limited to only offline
planning. It is able to render a graphical representation of joint
executions in the DOT graph description language, from which the diagrams
in Chapter \ref{ch:jointexec} were generated. This version also runs
on the Mozart platform version 1.3.2 or later. 
\item [{{PKnows:}}] This is our preliminary implementation of an epistemic
reasoning system using the techniques developed in Chapters \ref{ch:knowledge}
and \ref{ch:cknowledge}. It is implemented using SWI-Prolog to perform
symbolic manipulation, calling a modified version of the PDL prover
from the Tableaux Workbench suite \citep{abate07twb_desc} to handle
the resulting modal logic queries. 
\end{description}
