%
%  Top-Level File for PhD Thesis
%  Ryan Kelly
%

\include{preamble}

\begin{document}

\include{mathmacros}

\begin{titlepage}
\begin{center}
\ \\
\vspace{2cm}
{\bf\LARGE  ASYNCHRONOUS }\\ \vspace{0.5cm}
{\bf\LARGE  MULTI-AGENT REASONING }\\ \vspace{0.5cm}
{\bf\LARGE  IN THE SITUATION CALCULUS } \\
\vspace{3cm}
{\LARGE      Ryan Francis Kelly       }\\
\vspace{5cm}
{\em\large Submitted in total fulfilment of the requirements}\\ \vspace{0.1cm}
{\em\large        of the degree of Doctor of Philosophy     }\\
\vspace{0.5cm}
{\Large             October 2008        }\\
\vspace{2.5cm}
{\bf\large Department of Computer Science\\ and Software Engineering}\\ \vspace{0.5cm}
{\bf\Large        The University of Melbourne     }\\
\vspace{0.5cm}
\end{center}
\end{titlepage}

%Roman numeral numbering for initial section of thesis
\cleardoublepage
\pagenumbering{roman}

\include{abstract}

\chapter*{Declaration}
This is to certify that:
\begin{itemize}
\item[(i)] the thesis comprises only my original work towards the PhD except where indicated in the Preface,
\item[(ii)] due acknowledgement has been made in the text to all other material used,
\item[(iii)] the thesis is less than 100,000 words in length, exclusive of tables, maps, bibliographies, and appendices.
\end{itemize}
\vspace{3cm}
\rule{70mm}{0.1mm}\\
\emph{Ryan Francis Kelly}

\chapter*{Preface}
During the course of this research, a number of public presentations have been
 made which are based on the work presented in this thesis. They are listed
 here for reference.

\nobibliography*
\begin{itemize}
\item \bibentry{kelly06hlp_dps}
\item \bibentry{kelly07sc_persistence}
\item \bibentry{kelly07sc_know_obs}
\item \bibentry{kelly08complex_epistemic_modalities}
\end{itemize}


\chapter*{Acknowledgements} 

First, my heartfelt thanks to my supervisor, Adrian Pearce, for his support
and guidance throughout this research: for constantly reminding me to remember
the bigger picture; for encouraging me to explore any number of wild tangents; 
and for asking the deep questions about how they all join back together.
Thank you also for trusting me to work in my own time, in my own way, and for
always remaining confident that we would get there in the end.

To my advisory committee, Leon and James, thank you for providing the
clear heads needed to help keep the overall story in focus.  It seems like I
entered each of our progress meetings a little unsure about how all the pieces fit together, but left each with a renewed focus and a clear picture of the road ahead.

I must also acknowledge the generous support of the University of Melbourne
and the Australian Federal Government in providing my Australian Postgraduate
Award scholarship, without which this research would not have been possible.

To my fellow students and colleagues at the University of Melbourne -- particularly my office-mates Alan, Terence, Michelle, Arif, and my CSSEPG accomplices Alauddin, Kapil, Archana, Andrea, Macros -- thanks for making the whole postgrad
experience so rewarding. My thanks also to Sebastian Sardi\~na for his valuable feedback and many helpful discussions.

For my parents Don and Judy, my brother Matt, and my grandparents Val and
George, thank you for your unwavering support -- financial, physical and emotional -- throughout all of my studies.  Despite the constant jokes about me being ready
for retirement by the time I graduate (those \emph{were} just jokes...right?) I
have always known that I could count on your unconditional love and support.

Finally, and of course most importantly, to my wife Lauren: thank you
for your unending love and understanding; for your tolerance for the vagaries of
the PhD lifestyle; for keeping me firmly grounded in the real
world; and for simply being so excited about the whole endeavour.




% Contents and other lists are single-spaced
\singlespace
\cleardoublepage
\tableofcontents
\listoftables
\listoffigures
\listofalgorithms

% One-half line spaces for main body
% Also swicth to arabic page numbers
\cleardoublepage
\onehalfspace
\pagenumbering{arabic}

\include{intro}
\include{background}
\include{mindigolog}
\include{observations}
\include{jointexec}
\include{persistence}
\include{knowledge}
\include{cknowledge}
\include{conclusion}

% single-spacing for bibliography and appendices
\singlespace
\include{bibcmds}
\appendix
\include{proofs}
\include{implementation}
\include{cookingagents}
\end{document}

